### 2023-04-23 chapter2/chapter2.tex
Updated introduction outline
Finished BM analysis paragraph

\section{Introduction}
Here's where my introduction will go!
% I. IDRs, proteins without defined structures, are prevalent in eukaryotic proteins and play key roles in many processes
% II. Sequence-function relationships are unclear in IDRs
% III. Proteome-wide patterns of conserved features linked to function
%    A. Past studies have demonstrated evidence of constraint in specific proteins for specific properties
%    B. Additionally other studies have shown importance of short linear motifs in IDRs
%        1. Can be identified as short segments of conservation in otherwise highly diverged alignments of IDRs
%        2. However, these segments comprise a small fraction of disordered regions, so
%    C. Recent study first to demonstrate patterns of selection called evolutionary signatures in IDRs across entire proteome using a set of 82 molecular features
%    D. Furthermore, by clustering IDRs on by these evolutionary signatures, they identified groups of IDRs evolving under similar sets of constraints. Furthermore they identified functional associations in these clusters which can be used to generate hypotheses of function for previously uncharacterized proteins
% IV. Unclear if these patterns are specific to yeast or representative of IDR evolution across the tree of life
%    A. Expand analyses to another model organism
%        1. Drosophila is key model for genetic and development research with abundant genomic resources across many lineages in the radiation
%        2. Furthermore as multicellular organism with complex developmental pathways, the results would be particularly relevant for understanding the role of IDRs in complex signaling pathways
%        3. Also relevant for human health!
%    B. Results would have significant impact on interpretation
%        1. Conserved distributed properties are a yeast phenomena
%        2. Patterns of conservation are yeast phenomena
%        3. Global taxonomy of disordered regions similar to Pfam
% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
% V. By applying these approaches to set of IDRs identified in Drosophila genomes, we observe...

@@

% RESULTS
% IV. Brownian motion feature analyses
%     A. Having calculated features for all IDRs in our data set, we were curious the distribution contained any global structure which would allow the identification of classes with distinct biophysical or functional properties
%     A. Feature roots
%         1. Discuss differences in scaling and need for z-scores
%         2. Discussion of sample regions in figure? (need to add)
%     B. The phylogenetic contrasts algorithm also permits an estimate of the rate of feature evolution, which is effectively the variance of the distribution of features across all sequences after correcting for the phylogenetic relationships between them
%         1. PC1 correlates with rate
%         2. PC2 and PC3 show three major axes of unconstrained variation: Q-rich, G-rich, acidic

% V. BM-OU analyses
%    A. While the the BM model can the rates of change of features after accounting the the phylogenetic relationships between sequences, they do not directly test for conservation of features
%         1. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time
%         2. Instead, demonstrating evidence of conservation requires a comparison to a model where feature evolution is constrained.
%         3. A common choice is the Ornstein-Uhlenbeck (OU) model
%             i. Some explanation about how it works

>>>>

\section{Introduction}
Here's where my introduction will go!
% I. IDRs, proteins without defined structures, are prevalent in eukaryotic proteins and play key roles in many processes
% II. Sequence-function relationships are unclear in IDRs
% III. Proteome-wide patterns of conserved features linked to function
%    A. However, there is growing evidence that IDRs are instead constrained to conserve distributed properties such as ...
%       1. Because many sequences can yield similar a region with a specific charge or length, IDRs can rapidly diverge in terms of their sequences while still remaining
%       2. Additionally, in contrast to folded regions which must make precise contacts or satisfy specific geometric constraints, distributed features are robust to small changes at the sequence level. A mutation that slightly impairs function related to a specific feature regime is easily corrected by compensatory mutation elsewhere in the region. Together, these loose restrictions on the sequences of IDRs allow them to rapidly diverge despite preserving their functions.
%    B. Additionally other studies have shown importance of short linear motifs in IDRs
%        1. Can be identified as short segments of conservation in otherwise highly diverged alignments of IDRs
%        2. However, these segments comprise a small fraction of disordered regions, so
%    C. Recent study first to demonstrate patterns of selection in IDRs across an entire proteome using a set of 82 ``molecular features''
%    D. Furthermore, by clustering IDRs on by these ``evolutionary signatures,'' they identified groups of IDRs evolving under similar sets of constraints. Furthermore they identified functional associations in these clusters which can be used to generate hypotheses of function for previously uncharacterized proteins
% IV. However, unclear if these patterns are specific to yeast or representative of IDR evolution across the tree of life
%    A. Expand analyses to another model organism
%        1. Drosophila is key model for genetic and development research with abundant genomic resources across many lineages in the radiation
%        2. Furthermore as multicellular organism with complex developmental pathways, the results would be particularly relevant for understanding the role of IDRs in complex signaling pathways
%        3. Also relevant for human health!
%    B. Results would have significant impact on interpretation
%        1. Conserved distributed properties are a yeast phenomena
%        2. Patterns of conservation are yeast phenomena
%        3. Global taxonomy of disordered regions similar to Pfam
% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
% V. By applying these approaches to set of IDRs identified in Drosophila genomes, we observe...

@@

\subsection{IDRs have three axes of unconstrained variation}
Having calculated the features associated with the sequence segments composing each region in our data set, we then sought to determine if their distributions contained any global structure which would enable us to identify classes with distinct biophysical or functional properties. These distributions are generated by a complex underlying evolutionary process which reflects the combined effects of selection, drift, and mutation. However, to leverage a statistical framework to infer the properties of this process, we fit Brownian motion (BM) models to the features calculated from the segments in each region. BM is a simple model of evolution where continuous traits change through a series of small, undirected steps. Thus, the traits accumulate variation at a constant rate over time but do not on average deviate from their original values. BM models are therefore specified by two parameters: a rate, which describes the speed at which trait variation accumulates, and a root, which describes the ancestral trait value.

We then applied principal components analyses (PCAs) to visualize the major axes of variation of the root and rate parameters for each feature and region. A difficulty with a direct analysis of the parameter estimates, however, is PCAs are sensitive to differences in scaling between variables, and some features have dramatically different intrinsic scales. For example, many compositional features, like fraction\_S, are restricted to the interval [0, 1] whereas SCD is unbounded and can vary from negative to positive infinity. As a result, SCD is responsible for a significant fraction of the overall variance in both parameter distributions (Fig.~\ref{sfig:root_variance}-\ref{sfig:rate_variance}). Therefore, we first transformed the parameters associated with each feature into z-scores relative to the proteome-wide distributions.

The first two principal components of the root distributions show little overall structure, though there is a slight enrichment of regions along two axes that correlate with acidic and polar features, respectively (Fig.~\ref{fig:root}). Likewise, the projections of the rates onto the first two components are largely distributed along the first principal component (Fig.~\ref{fig:rate}A). This and the variable amounts of sequence divergence in the regions led us to suspect the first principal component was a measure of the overall rate of sequence evolution. Plotting the first principal component against the sum of the average amino acid and indel rates as measured by substitution models revealed a strong association (Fig.~\ref{fig:rate}B). We then projected the rates along second and third principal components to determine if these higher order components contained any additional structure. The resulting distribution is roughly triangular and contains three major axes of variation, corresponding to rapid changes in the regions' proportions of acidic, glutamine, and glycine residues (Fig.~\ref{fig:rate}C-D). Inspection of regions selected along these axes confirmed the high rates of evolution of these features (Fig.~\ref{fig:rate}E-G).

% RESULTS
% V. BM-OU analyses
%    A. While the the BM model can the rates of change of features after accounting the the phylogenetic relationships between sequences, they do not directly test for conservation of features
%         1. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time
%         2. Instead, demonstrating evidence of conservation requires a comparison to a model where feature evolution is constrained.
%         3. A common choice is the Ornstein-Uhlenbeck (OU) model
%             i. Some explanation about how it works
